\section*{Abstract}

This thesis focuses on game development for Android, an operating system for mobile phones developed by Google. The market for Android applications is currently growing rapidly, and there are currently many games available. The purpose of this thesis is to find out how to develop a competitive Tower Defense game on this expanding market.

Tower Defense is not a new game concept. There are several games of this type with different  themes and gameplay already on the market. In order to provide a solid foundation from which a new Tower Defense game can be developed, many of the existing games were examined thoroughly to find desirable features. In addition to this, interviews with people that play Tower Defense games frequently were conducted, making sure no important features were left out.

The conclusion of the research was that in order to make the game competitive on the Android Market, the gameplay had to be unique. The features of the mobile device were used as tools to achieve this goal. Smartphones primarily provide two features that were taken into consideration when developing Eskimo Tower Defense; the touchscreen and the accelerometer. The touchscreen provides a new interaction method in comparison with similar games on a computer platform. However, this alone did not suffice to distinguish a Tower Defense game on the Android platform. Eskimo Tower Defense utilizes the accelerometer to give players a feeling of being able to control the game, making them more physically involved in the gameplay.