\section*{Abstract}

This thesis focuses on game development for Android, an operating system for mobile phones developed by Google. The market for Android applications is currently growing rapidly, and there are currently many games available. The purpose of this thesis is to find out how to develop a competitive Tower Defense game on this expanding market.

Tower Defense is not a new game concept. There are several games of this type with different  themes and gameplay already on the market. In order to provide a solid foundation from which a new Tower Defense game can be developed, many of the existing games were examined thoroughly to find desirable features. In addition to this, interviews with people that play Tower Defense games frequently were conducted, making sure no important features were left out.

The conclusion of the research was that in order to make the game competitive on the Android Market, the gameplay had to be unique. The features of the mobile device were used as tools to achieve this goal. Smartphones primarily provide two features that were taken into consideration when developing Eskimo Tower Defense; the touchscreen and the accelerometer. The touchscreen provides a new interaction method in comparison with similar games on a computer platform. However, this alone did not suffice to distinguish a Tower Defense game on the Android platform. Eskimo Tower Defense utilizes the accelerometer to give players a feeling of being able to control the game, making them more physically involved in the gameplay.

\section*{Sammanfattning}

Den här rapporten fokuserar på spelutveckling för Android, ett operativsystem för mobiltelefoner utvecklat av Google. Marknaden för applikationer till Android växer kraftigt och en stor del av dessa är spel. Syftet med det här arbetet är att utreda hur en konkurrenskraftigt Tower Defense-spel kan utvecklas för den här växande marknaden.

Tower Defense är ett välkänt spelkoncept. Det finns ett flertal spel av den typen med olika teman och spelupplevelse på marknaden. För att skapa en stabil grund för utveckling av ett Tower Defense-spel undersöktes flera existerande Tower Defense-spel utförligt. Utöver detta genomfördes intervjuer med personer som spelar Tower Defense-spel regelbundet, för att tillse att inga viktiga funktioner utelämnades.

Slutsatserna från undersökningen visade att för att ett spel på Android-marknaden ska vara konkurrenskraftigt så behöver det tillhandahålla en unik spelkänsla. Funktionerna hos mobiltelefonerna användes som verktyg för att uppnå detta mål. De funktioner som främst tillämpades under utvecklingen av Eskimo Tower Defense var pekskärmen och accelerometern. Pekskärmen möjliggör interaktion på ett sätt som inte används i vanliga datorspel inom samma genre. Värden från accelerometersensorer tillämpas för att få användaren att känna sig mer involverad under spelets gång.